---
title: "Import data into Epigraf"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{import}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  message = FALSE, 
  warning = FALSE
)

library(tidyverse)
library(httr)
library(rpigraf)

api_setup("http://localhost", "devel")

```

Epigraf stores data in a format called Relational Article Model (RAM).

RAM is a model to represent documents in tables. 
It consists of the following tables:

- Projects: Each article is assigned to a project.
- Articles: Each article contains sections.
- Sections: Each section contains items.
- Items: Items contain the article content (text, references to categories etc.)
- Links: Annotations are stored as links between tags in item content and categories.
- Footnotes: Textual annotations are stored in the footnotes table.
- Categories: Vocabularies used for tagging items and annotating text.

Each row in a table is identified by an IRI (Internationalized Resource Identifier),
which is a worldwide unique identificator. 
Understanding the IRI concept of the RAM model is crucial for data transfers:

- Add new data: When you import data with IRIs not yet present in the database,
  new rows will be added.
- Update data:  When you import data with IRIs that already exist, 
  the existing rows will be updated.
  
Each IRI used in the RAM consists of the server name (e.g. "https://epigraf.uni-muenster.de/")
and an IRI path (e.g. "properties/genres/western"). 

The IRI path is stored in the database tables and consists of three elements:
- Table: One of projects, articles, sections, items, or categories.
- Type: The row type. For example, you can distinguish different category systems by their type.
- Fragment: An alphanumeric value used to identify the row in the table.


## Download data

### Fetch data

First, you fetch RAM data from Epigraf. 
The data contains rows from the RAM tables (projects, articles, sections etc.) below each other.

```{r eval=FALSE}
epi <- api_fetch("articles", db = "epi_movies")

```

### Map RAM data to articles and their annotations

Second, although you can directly work with the RAM data, we suggest you let rpigraf distill flat data frames for you.
You can start from two perspectives:

- Get articles and, optionally, join nested content with `distill_articles()`.
- Get properties and, optionally, join annotations from the articles with `distill_properties()`.

#### Get article content

The distill_articles() method pulls article data from the RAM data and joins sections, items and optionally properties used in the items.
In its most simple form, you get a list of articles with selected columns:

```{r eval=FALSE}
distill_articles(epi, c("signature", "name"))
```

To join item data, provide an item type and the columns to extract from the items.
This results in a data frame with a row for each item (and the respective article columns).

```{r eval=FALSE}
distill_articles(epi, c("name","signature"), item.type = "text", item.cols = "content")

```

If properties are used in the items, add a parameter that defines which fields to extract from the joined property:

```{r eval=FALSE}
distill_articles(epi, c("signature", "name"), item.type = "categories", property.cols = "lemma")

```

#### Get properties and annotations

To get the properties used in the articles (i.e. properties that were fetched and, thus, are contained in the RAM data),
you use `distill_properties()` with RAM data in the first parameter and a property type in the second parameter.
For hierarchical properties, the function will return the property tree.

```{r eval=FALSE}
distill_properties(epi, "categories")
```

To get the annotations, including coded segments, set the `annos`-parameter to TRUE:

```{r eval=FALSE}
distill_properties(epi, "categories", annos = TRUE)
distill_properties(epi, "annotations", annos = TRUE)

```

