---
title: "Import data into Epigraf"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{import}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  message = FALSE, 
  warning = FALSE
)

library(tidyverse)
library(httr)
library(rpigraf)

api_setup("http://localhost", "devel")

```

Epigraf expects data in a format called Relational Article Model (RAM).

RAM is a model to represent documents in tables. 
It consists of the following tables:

- Projects: Each article is assigned to a project.
- Articles: Each article contains sections.
- Sections: Each section contains items.
- Items: Items contain the article content (text, references to categories etc.)
- Categories: Vocabularies used for tagging items and annotating text.

Each row in a table is identified by an IRI (Internationalized Resource Identifier),
which is a worldwide unique identificator. 
Understanding the IRI concept of the RAM model is crucial for data transfers:

- Add new data: When you import data with IRIs not yet present in the database,
  new rows will be added.
- Update data:  When you import data with IRIs that already exist, 
  the existing rows will be updated.
  
Each IRI used in the RAM consists of the server name (e.g. "https://epigraf.uni-muenster.de/")
and an IRI path (e.g. "properties/genres/western"). 

The IRI path is stored in the database tables and consists of three elements:
- Table: One of projects, articles, sections, items, or categories.
- Type: The row type. For example, you can distinguish different category systems by their type.
- Fragment: An alphanumeric value used to identify the row in the table.


## Upload data

We use the following simple table as an example:

```{r}

ds <- tribble(
  ~case, ~title,       ~genre,     ~text,
  "011", "Westworld",   "Western", "In a futuristic amusement park, androids populate themed worlds like the Wild West.",
  "012", "Yellowstone", "Western", "Ranch owner John Dutton battles to protect his family's massive Montana cattle ranch.",
  "013", "Once Upon",   "Western", "A mysterious harmonica-playing stranger teams up with a bandit."
)
```

### Map data to the RAM

The craft methods guide your through the mapping process.

Project: As we don't have data about the project in the source data frame, we fill in some values.
The project is named "Movies".
The type defaults to "default" (optional, if you omit it, "default" is used anyways)

```{r}
ds <- craft_projects(ds, fill = c("type" = "default", "fragment" = "movies",  "name" = "Movies", "signature"="movies"))
```


Article data is taken from the columns: case, title. 
The type defaults to "default".
```{r}
ds <- craft_articles(ds, fill = c("type" = "default"), cols=c("fragment" = "case", "signature" = "case", "name" = "title"))
```


The text section is named "Abstract".
We omit the fragment, it is automatically added and derived from the article fragment.
You need a fragment, if you have multiple sections of the same type.
```{r}
ds <- craft_sections(ds, fill = c("type" = "text", "name" = "Abstract"))
```

The text item data is taken from the columns: text.
The type is fixed to "text".
We omit the fragment, it is automatically added and derived from the section fragment.
You need a fragment, if you have multiple items of the same type.

```{r}
ds <- craft_items(ds, fill = c("type" = "text"), cols = c("content" = "text"))
```

The categories section is named "Genres".
The type is fixed to "categories".

```{r}
ds <- craft_sections(ds, fill = c("type" = "categories", "name" = "Genres"))
```

The categories property content is taken from the genre column.
By crafting a property, the column .property is added to the data frame.
The categories item content is taken from this column.

```{r}
ds <- craft_properties(ds, fill = c("type" = "categories"), cols = c("fragment"="genre", "lemma" = "genre"))
ds <- craft_items(ds, fill = c("type" = "categories"), cols = c("properties_id" = ".property"))
```



### Patch data into the database

Finally, compile the stacked ram rows to get the RAM data frame.
```{r}
epi <- ram_compile(ds)
```


The RAM data frame is ready to be patched (=uploaded) into Epigraf.
```{r eval=FALSE}
api_patch(epi, db = "epi_movies")
```


## Download data

### Fetch data

```{r eval=FALSE}
epi <- api_fetch("articles", db = "epi_movies")
```

### Map RAM data to articles and their annotations

Get articles with some item data
```{r eval=FALSE}
distill_articles(epi, c("signature", "name"), item.type = "categories", property.cols = "lemma")

distill_articles(epi, c("name","signature"), item.type = "text", item.cols = "content")
```


Get the properties used in the articles.
```{r eval=FALSE}
distill_properties(epi, "categories")
```


Get the annotations, if any, for a property type.
```{r eval=FALSE}
distill_properties(epi, "categories", annos = TRUE)
distill_properties(epi, "annotations", annos = TRUE)
```
